---
title: "Treatment Selection Simulations"
output: html_document
---

```{r, message = F, warning = F }
library(tidyverse)
library(lme4)
library(kableExtra)
source("sim.R")
```

# Minimum Sample Size vs PCS

How many samples do we need before we achieve a minimum 90% PCS

### Variables

- Period Length: 1 - 50
- Effect of Treatment B
- Standard Deviation of Treaments

### Design Parameters

- Sampling Frequency: 1
- Treatment Order: A, B, B, A
- Number of Blocks: 2

### Treatment Parameters

- Effect Size: 5
- Effect of Treatment A: -5
- Run-in of A: $\tau_a = 2$
- Carryover of A: $\gamma_a = 2$
- Run-in of B: $\tau_b = 2$
- Carryover of B: $\gamma_b = 2$

### Outcome Parameters

- $\mu_b = 100$
- $\sigma_b^2 = 1$

```{r}
N = 100
samp.candidates = 1:200
b.effs = c(0, -1, -2, -3, -4)
sd.vec = seq(2, 10, by = 2)

pcs.v.minsamp.mu.b = 100
pcs.v.minsamp.sd.b = 1
pcs.v.minsamp.sd.o = 0.01
pcs.v.minsamp.order = c("A", "B", "C")
pcs.v.minsamp.s.freq = 1  
pcs.v.minsamp.n.blocks = 2
pcs.v.minsamp.results = NULL

for (beff in b.effs) {
  for (s in sd.vec) {
    
    prog.txt = paste0("Simulations for B effect ", beff, " with treatment sd ", s)
    print(prog.txt)
    
     trt.A = list(
      name = "A",
      effect = -5,
      sd = s,
      run = 2,
      carry = 2
      )
     
     trt.B = list(
      name = "B",
      effect = beff,
      sd = s,
      run = 2,
      carry = 2
      )
     
     trt.C = list(
      name = "C",
      effect = 0,
      sd = 0,
      run = 1,
      carry = 1
      )
     
     pcs.v.minsamp.trts = list(
      testing = c("A", "B", "C"),
      A = trt.A,
      B = trt.B,
      C = trt.C
      )
     
     for (cand.p in samp.candidates) {
       
      pcs.vec = integer(N)
      
      for (i in 1:N) {
        
        sim.data = simulate.trial(pcs.v.minsamp.trts, 
                                  mu.b = pcs.v.minsamp.mu.b, 
                                  sd.b = pcs.v.minsamp.sd.b,
                                  sd.o = pcs.v.minsamp.sd.o,
                                  order = pcs.v.minsamp.order, 
                                  s.freq = pcs.v.minsamp.s.freq, 
                                  p.length = cand.p, 
                                  n.blocks = pcs.v.minsamp.n.blocks,
                                  baseline.type = "markov")
        
        sim.data = sim.data %>% 
          mutate(
            treatment = factor(treatment, levels = c("C", "A", "B"))
          )
        
        # Run the model on the data
        sim.model = lm(obs ~ factor(treatment) + factor(block), data = sim.data)
        
        coeff.tbl = summary(sim.model)$coefficients
        
        pcs.vec[i] = ifelse(abs(coeff.tbl[2,1]) > abs(coeff.tbl[3,1]), 1, 0)
      }
      
      if (mean(pcs.vec) > .9) { 
        print(paste0("Stopped at N = ", cand.p))
        break 
        }
      
     }
     # Record the results of these parameters
     res = tibble(
       eff.diff = -5 - beff,
       trt.noise = s,
       min.samp = cand.p
       )
     
     pcs.v.minsamp.results = bind_rows(pcs.v.minsamp.results, res)
    }
  }


pcs.v.minsamp.results = pcs.v.minsamp.results %>% 
  mutate(
    `Effect Difference` = factor(eff.diff)
  ) 

pcs.plot = pcs.v.minsamp.results %>% 
  ggplot(aes(x = trt.noise, y = min.samp, color = `Effect Difference`, group = `Effect Difference`)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Minimum period length needed to achieve 90% PCS (maximal)",
    x = "Treatment noise",
    y = "Minimum period length to reach 90%"
  ) + 
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
    panel.background = element_rect(fill = "white")
  )

pcs.plot
```

```{r}
write.csv(pcs.v.minsamp.results, "../img/pcs-v-minsamp-results.csv")
```

### Achieving Treatment Window

```{r}
N = 100
samp.candidates = 1:200
window.sizes = c(5, 4, 3, 2, 1)
sd.vec = seq(2, 10, by = 2)

pcs.window.v.minsamp.mu.b = 100
pcs.window.v.minsamp.sd.b = 1
pcs.window.v.minsamp.sd.o = 0.01
pcs.window.v.minsamp.order = c("A", "B", "C", "D")
pcs.window.v.minsamp.s.freq = 1  
pcs.window.v.minsamp.n.blocks = 2
pcs.window.v.minsamp.results = NULL

for (window in window.sizes) {
  for (s in sd.vec) {
    
    prog.txt = paste0("Simulations for Window size ", window, " with treatment sd ", s)
    print(prog.txt)
    
     trt.A = list(
      name = "A",
      effect = -5,
      sd = s,
      run = 2,
      carry = 2
      )
     
     trt.B = list(
      name = "B",
      effect = -5 + window,
      sd = s,
      run = 2,
      carry = 2
      )
     
     trt.C = list(
      name = "C",
      effect = -5 - window,
      sd = s,
      run = 2,
      carry = 2
      )
     
     trt.D = list(
      name = "D",
      effect = 0,
      sd = 0,
      run = 0,
      carry = 0
      )
     
     pcs.window.v.minsamp.trts = list(
      testing = c("A", "B", "C", "D"),
      A = trt.A,
      B = trt.B,
      C = trt.C,
      D = trt.D
      )
     
     for (cand.p in samp.candidates) {
       
      pcs.vec = integer(N)
      
      for (i in 1:N) {
        
        sim.data = simulate.trial(pcs.window.v.minsamp.trts, 
                                  mu.b = pcs.window.v.minsamp.mu.b, 
                                  sd.b = pcs.window.v.minsamp.sd.b,
                                  sd.o = pcs.window.v.minsamp.sd.o,
                                  order = pcs.window.v.minsamp.order, 
                                  s.freq = pcs.window.v.minsamp.s.freq, 
                                  p.length = cand.p, 
                                  n.blocks = pcs.window.v.minsamp.n.blocks,
                                  baseline.type = "markov")
        
        sim.data = sim.data %>% 
          mutate(
            treatment = factor(treatment, levels = c("D", "A", "B", "C"))
          )
        
        # Run the model on the data
        sim.model = lm(obs ~ factor(treatment) + factor(block), data = sim.data)
        
        coeff.tbl = summary(sim.model)$coefficients
        
        # Gather coefficients into a vector for comparison
        coeffs = coeff.tbl[2:4,1]
        deviations =  abs(coeffs + 5) # We know the optimal treatment is A
        
        # Check the index of the treatment that most closely achieves the true effect
        best.idx = which(deviations == min(deviations))
        
        pcs.vec[i] = ifelse(best.idx == 1, 1, 0)
      }
      
      if (mean(pcs.vec) > .9) { 
        print(paste0("Stopped at N = ", cand.p))
        break 
        }
      
     }
     # Record the results of these parameters
     res = tibble(
       window.size = window,
       trt.noise = s,
       min.samp = cand.p
       )
     
     pcs.window.v.minsamp.results = bind_rows(pcs.window.v.minsamp.results, res)
    }
  }


pcs.window.v.minsamp.results = pcs.window.v.minsamp.results %>% 
  mutate(
    `Window Size` = factor(window.size)
  ) 

pcs.plot2 = pcs.window.v.minsamp.results %>% 
  ggplot(aes(x = trt.noise, y = min.samp, color = `Window Size`, group = `Window Size`)) +
  geom_point() +
  geom_line() +
  labs(
    title = "Minimum period length needed to achieve 90% PCS (window)",
    x = "Treatment noise",
    y = "Minimum period length needed to reach 90%"
  ) + 
  theme(
    plot.title = element_text(hjust = 0.5),
    legend.position = "bottom",
    panel.background = element_rect(fill = "white")
  )

pcs.plot2
write.csv(pcs.window.v.minsamp.results, "../img/pcs2-v-minsamp-results.csv")
```